---
title: "Land Matrix monitoring"
author: "Accountability"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  include = FALSE,
  comment = ''
)
```

# Notes

## What is it for?

Pull data from Land Matrix (using GraphQL) and apply a few filters + tests to assess data availability and quality.

## How do I use it?

- [Add GitHub repo for the R script]
- For more detail on how the filters work, use the documentation (available on the GitHub)
- Use the comments in the code to edit the filters if needed (always use "custom" parameters and not "default" parameters)

…

---

# Pull and filter data from landmatrix.org

```{r Load libraries}
# Load libraries
library(ghql)       # GraphQL client
library(jsonlite)   # Work with JSON objects in R
library(dplyr)
library(tibble)
library(knitr)      # Render nice tables in markdown
library(DT)         # Render interactive tables
library(ggplot2)
library(visdat)
library(naniar)
```

``` {r globalVar}
# ========== GLOBAL VARIABLES ==========
last_modification_limit <- 2021;  # Must be a year. If older, the deal has to be updated.
```

```{r Init client}
########## INIT GRAPHQL CLIENT ##########

# Link to the Land Matrix GraphQL API
link <- 'https://landmatrix.org/graphql/'

# Create a new graphqlClient connection object
conn <- GraphqlClient$new(url = link)

```

```{r Filters, echo = TRUE}
########## LAND MATRIX VGGTS FILTERS ##########
# WARNING → trailing commas at the end of the last filter make the query fail

filters <- list(
  # Deals
  #list( field = "id", operation = "IN", value = c(14, 15, 16, 18, 20) ),
  
  # Region
  #list( field = "country.region.id", operation = "EQ", value = 142 ),
  
  # Country
  list( field = "country.id", operation = "EQ", value = 120 ),
  
  # Deal size → at least 200 ha
  list( field = "deal_size", operation = "GE", value = 200 ),
  
  # Negotiation status → Concluded only (oral agreement, contract signed or change of ownership)
  list( field = "current_negotiation_status", 
        operation = "IN", 
        value = c("ORAL_AGREEMENT", "CONTRACT_SIGNED", "CHANGE_OF_OWNERSHIP") ),
  
  # Nature of the deal → outright purchase, lease, concession, exploitation permit, include unknown
  list( field = "nature_of_deal", 
        operation = "CONTAINED_BY", 
        value = c("OUTRIGHT_PURCHASE", "LEASE", "CONCESSION", "EXPLOITATION_PERMIT"),
        allow_null = TRUE),
  
  # Year of initiation → 2000 or later, include unknown
  list( field = "initiation_year", operation = "GE", value = 2000, allow_null = TRUE ),
  
  # Intention of investment → agriculture, forestry
  list( field = "current_intention_of_investment", 
        operation = "OVERLAP", 
        value = c(
          "BIOFUELS", "FOOD_CROPS", "FODDER", "LIVESTOCK", "NON_FOOD_AGRICULTURE", "AGRICULTURE_UNSPECIFIED",
          "TIMBER_PLANTATION", "FOREST_LOGGING", "CARBON", "FORESTRY_UNSPECIFIED"
          )),
  
  # Scope → transnational
  list( field = "transnational", operation = "EQ", value = TRUE )
  
)

# Deal 8245 missing → in the platform, nature of deal = other is included by default in the platform 

```

```{r Fields and variables}
# GraphQL query string with variables and Land Matrix fields to query
query_string <- '
  query(
    $sort: String
    $limit: Int
    $filters: [Filter]
    ){
    deals(
      sort: $sort
      limit: $limit
      filters: $filters
    ){
      id
      country {id, name}
      
      deal_size
      current_negotiation_status
      nature_of_deal
      initiation_year
      current_intention_of_investment
      created_at
      modified_at
      fully_updated
      fully_updated_at
      
      recognition_status
      recognition_status_comment
      
      displacement_of_people
      displaced_people
      displaced_households
      displaced_people_from_community_land
      displaced_people_within_community_land
      displaced_households_from_fields
      displaced_people_on_completion
      displacement_of_people_comment
      
      community_consultation
      community_consultation_comment
      
      community_reaction
      community_reaction_comment
      
      land_conflicts
      land_conflicts_comment
      
      contract_farming
      on_the_lease_state
      on_the_lease
      off_the_lease_state
      off_the_lease
      contract_farming_comment
      
      promised_benefits
      promised_benefits_comment
      materialized_benefits
      materialized_benefits_comment
      
      negative_impacts
      negative_impacts_comment
      
      promised_compensation
      
      promised_compensation
      received_compensation
      
      gender_related_information
      
      purchase_price
      purchase_price_currency { id name }
      purchase_price_type
      purchase_price_area
      purchase_price_comment
      
      annual_leasing_fee
      annual_leasing_fee_currency { id name }
      annual_leasing_fee_type
      annual_leasing_fee_area
      annual_leasing_fee_comment
      
      presence_of_organizations
      
      involved_actors
    }
  }
'

# Variables
variables <- list(
  sort = "id",
  limit = 100000,
  filters = filters
)
```

```{r Query results}

# Use the ghql class to define the query object
query <- Query$new()$query('link', query_string)

# Make the request, pass the query and then the variables. Then convert
# the raw object to a structure JSON object.
result <- conn$exec(query$link, variables = variables) %>%
  fromJSON(flatten = T)

# Convert the JSON data into a tibble
deals_data <- result$data$deals %>%
  as_tibble()

# All empty strings to NA
deals_data <- deals_data %>%
  mutate_if(is.character, list(~na_if(.,"")))

deals_data

```

Inspection of the schema sent to Land Matrix:

```{r echo=FALSE, include=TRUE}
query$link
```

List of filters:
```{r include=TRUE}
# Create vectors to store data in and turn into filters later
c_fields <- c()
c_operations <- c()
c_values <- c()

# Loop over the filters to create tibble rows
for (f in filters) {
  c_fields <- c(c_fields, f$field)
  c_operations <- c(c_operations, f$operation)
  c_values <- c(c_values, paste(f$value, collapse=", ")) # paste() + collapse argument turns a vector into a string so all values fit in one column
}

# Write all filters in a tibble
filters_table <- tibble(c_fields, c_operations, c_values) %>%
  rename(Field = c_fields, Operation = c_operations, Value = c_values)
kable(filters_table)

```

---

# Results

```{r include=TRUE}
deals_data %>%
  select(id, country.name, created_at, modified_at, fully_updated_at) %>%
  mutate(
    created_at = format(as.POSIXct(created_at), format="%Y-%m-%d"),
    modified_at = format(as.POSIXct(modified_at), format="%Y-%m-%d"),
    fully_updated_at = format(as.POSIXct(fully_updated_at), format="%Y-%m-%d"),
  ) %>%
  datatable(fillContainer = TRUE, 
            options = list(scrollY = "380px")
            )
```

---

# Metrics

## Number of deals, number of countries

## Most recent data

```{r include = TRUE}
deals_data %>%
  select(id, country.name, created_at, modified_at, fully_updated, fully_updated_at) %>%
  mutate(
    created_at = format(as.POSIXct(created_at), format="%Y-%m-%d"),
    modified_at = format(as.POSIXct(modified_at), format="%Y-%m-%d"),
    fully_updated_at = format(as.POSIXct(fully_updated_at), format="%Y-%m-%d"),
  ) %>%
  datatable(fillContainer = TRUE, 
            options = list(scrollY = "380px")
            )
```

```{r include = TRUE, fig.height=4, fig.width=4}
# fig.height = 10 (regional level) or 24 with 4 columns (global level) or 50 with 4 columns (all countries)
# fig.width = 12 (down to 4 for one country only)

deals_data %>%
  select(country.name, created_at) %>%
  mutate(
    creation_year = format(as.POSIXct(created_at), format="%Y")
  ) %>%
  ggplot(aes(x=creation_year)) + geom_bar() +
  facet_wrap(~country.name,
             ncol = 3,
             ) +
  labs(
    x="Creation date",
    y="Number of deals"
  ) +
  theme_bw()
```

## Deals that must be fully updated

This is a list of the deals that have been last modified before `r last_modification_limit`.
They should be updated (we want data with a full update less than 2 years old).

```{r include = TRUE}

# Get all deals older than last_modification_limit (modified_at)
deals_data %>%
  select(id, country.name, fully_updated_at) %>%
  mutate(year = format(as.POSIXct(fully_updated_at), format="%Y")) %>%
  mutate(last_full_update = format(as.POSIXct(fully_updated_at), format="%Y-%m-%d")) %>%
  filter(year < last_modification_limit) %>%
  arrange(year) %>%
  select(id, country.name, last_full_update) %>%
  datatable(fillContainer = TRUE, 
            options = list(scrollY = "380px")
            )

```

---

# Transparency

## NAs

Number of deals assessed: `r nrow(deals_data)`.

```{r include = TRUE, fig.height = 16, fig.width = 8}
deals_na <- deals_data %>%
  summarise_all(funs(sum(is.na(.))))

deals_na <- as.data.frame(t(deals_na))

deals_na %>% 
  mutate("Percentage of NA" = round(V1/nrow(deals_data)*100, digits = 1)) %>%
  rename("Number of NA" = V1) %>%
  kable()

deals_data %>%
  gg_miss_var(show_pct = TRUE)
```

```{r include = TRUE, fig.height = 6, fig.width = 8}

# Fields 1 to 14
deals_data %>%
  select(
         
     # Variable 1
    recognition_status, recognition_status_comment,
    
    # Variable 2
    displacement_of_people, displaced_people, displaced_households, displaced_people_from_community_land,
    displaced_people_within_community_land, displaced_households_from_fields, displaced_people_on_completion,
    displacement_of_people_comment,
    
    # Variable 3
    community_consultation, community_consultation_comment,
    
    # Variable 4
    community_reaction, community_reaction_comment,
    
    ) %>%
  vis_miss()

# Fields 15 to 28
deals_data %>%
  select(

     # Variable 5
    land_conflicts, land_conflicts_comment,
    
    # Variable 6
    contract_farming, on_the_lease_state, on_the_lease, off_the_lease_state, off_the_lease, contract_farming_comment,
    
    # Variables 7 and 8
    promised_benefits, promised_benefits_comment,
    materialized_benefits, materialized_benefits_comment,
    
    # Variable 9
    negative_impacts, negative_impacts_comment,
         
    ) %>%
  vis_miss()

# Fields 29 to 42
deals_data %>%
  select(

    # Variables 10 and 11
    promised_compensation,
    
    # Variable 11
    received_compensation,
    
    # Variable 12
    gender_related_information,
    
    # Variable 13
    purchase_price, purchase_price_currency.name, purchase_price_type, purchase_price_area, purchase_price_comment,
    
    # Variable 14
    annual_leasing_fee, annual_leasing_fee_currency.name, annual_leasing_fee_type, annual_leasing_fee_comment,
    
    # Variable 15
    presence_of_organizations,
    
    # Variable 16
    involved_actors
         
    ) %>%
  vis_miss()

```

```{r include = FALSE}
deals_data %>%
  select(id,
         
    # Variable 1
    recognition_status, recognition_status_comment,
    
    # Variable 2
    displacement_of_people, displaced_people, displaced_households, displaced_people_from_community_land,
    displaced_people_within_community_land, displaced_households_from_fields, displaced_people_on_completion,
    displacement_of_people_comment,
    
    # Variable 3
    community_consultation, community_consultation_comment,
    
    # Variable 4
    community_reaction, community_reaction_comment,
    
    # Variable 5
    land_conflicts, land_conflicts_comment,
    
    # Variable 6
    contract_farming, on_the_lease_state, on_the_lease, off_the_lease_state, off_the_lease, contract_farming_comment,
    
    # Variables 7 and 8
    promised_benefits, promised_benefits_comment,
    materialized_benefits, materialized_benefits_comment,
    
    # Variable 9
    negative_impacts, negative_impacts_comment,
    
    # Variables 10 and 11
    promised_compensation,
    
    # Variable 11
    received_compensation,
    
    # Variable 12
    gender_related_information,
    
    # Variable 13
    purchase_price, purchase_price_currency.name, purchase_price_type, purchase_price_area, purchase_price_comment,
    
    # Variable 14
    annual_leasing_fee, annual_leasing_fee_currency.name, annual_leasing_fee_type, annual_leasing_fee_comment,
    
    # Variable 15
    presence_of_organizations,
    
    # Variable 16
    involved_actors
    
  ) %>%
  datatable(fillContainer = TRUE, 
            options = list(scrollY = "380px")
            )
```

## Transparency score

Calculate the transparency score based on the VGGTs variables. The Land Matrix variables taken into account in the VGGTs are the following:

- recognition_status
- recognition_status_comment
- displacement_of_people
- displaced_people
- displaced_households
- displaced_people_from_community_land
- displaced_people_within_community_land
- displaced_households_from_fields
- displaced_people_on_completion
- displacement_of_people_comment
- community_consultation
- community_consultation_comment
- community_reaction
- community_reaction_comment
- land_conflicts
- land_conflicts_comment
- contract_farming
- on_the_lease_state
- on_the_lease
- off_the_lease_state
- off_the_lease
- off_the_lease
- promised_benefits
- promised_benefits_comment
- materialized_benefits
- materialized_benefits_comment
- negative_impacts
- negative_impacts_comment
- promised_compensation
- received_compensation
- gender_related_information
- purchase_price
- purchase_price_currency
- purchase_price_type
- purchase_price_area
- purchase_price_comment
- annual_leasing_fee
- annual_leasing_fee_currency
- annual_leasing_fee_type
- annual_leasing_fee_area
- annual_leasing_fee_comment
- presence_of_organizations
- involved_actors

---

# Retention tests

## Apply the retention tests

## Deals overview

- Deal ID
- Deal country
- Deal region
- Transparency score
- Retention test result
- Reason for retention test failure

## Countries overview

- Country
- Region
- Transparency score
- Number of deals
- Number of retained deals

---

# Missing data assessment

Display chapters with missing data (like in the original script).
Same for articles.

---