---
title: "Land Matrix monitoring"
author: "Accountability"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  include = FALSE,
  comment = ''
)
```

# Notes

## What is it for?

Pull data from Land Matrix (using GraphQL) and apply a few filters + tests to assess data availability and quality.

## How do I use it?

- [Add GitHub repo for the R script]
- For more detail on how the filters work, use the documentation (available on the GitHub)
- Use the comments in the code to edit the filters if needed (always use "custom" parameters and not "default" parameters)

…

---

# Pull and filter data from landmatrix.org

```{r Load libraries}
# Load libraries
library(ghql)       # GraphQL client
library(jsonlite)   # Work with JSON objects in R
library(dplyr)
library(tibble)
library(knitr)      # Render nice tables in markdown
library(DT)         # Render interactive tables
```

```{r Init client}
########## INIT GRAPHQL CLIENT ##########

# Link to the Land Matrix GraphQL API
link <- 'https://landmatrix.org/graphql/'

# Create a new graphqlClient connection object
conn <- GraphqlClient$new(url = link)

```

```{r Filters, echo = TRUE}
# Land Matrix filters
filters <- list(
  #list( field = "id", operation = "IN", value = c(14, 15, 16, 18, 20) ),
  list( field = "country.region.id", operation = "EQ", value = 2 ),
  list( field = "country.id", operation = "EQ", value = 710 ),
  list( field = "initiation_year", operation = "GE", value = 2000, allow_null = TRUE )
)
```

```{r Fields and variables}
# GraphQL query string with variables and Land Matrix fields to query
query_string <- '
  query(
    $sort: String
    $limit: Int
    $filters: [Filter]
    ){
    deals(
      sort: $sort
      limit: $limit
      filters: $filters
    ){
      id
      country {id, name}
      deal_size
      initiation_year
    }
  }
'

# Variables
variables <- list(
  sort = "id",
  limit = 6,
  filters = filters
)
```

```{r Query results}

# Use the ghql class to define the query object
query <- Query$new()$query('link', query_string)

# Make the request, pass the query and then the variables. Then convert
# the raw object to a structure JSON object.
result <- conn$exec(query$link, variables = variables) %>%
  fromJSON(flatten = T)

# Convert the JSON data into a tibble
deals_data <- result$data$deals %>%
  as_tibble()
deals_data

```

Inspect the schema:

```{r echo=FALSE, include=TRUE}
query$link
```

List of filters:
```{r include=TRUE}
# Create vectors to store data in and turn into filters later
c_fields <- c()
c_operations <- c()
c_values <- c()

# Loop over the filters to create tibble rows
for (f in filters) {
  c_fields <- c(c_fields, f$field)
  c_operations <- c(c_operations, f$operation)
  c_values <- c(c_values, f$value)
}

# Write all filters in a tibble
filters_table <- tibble(c_fields, c_operations, c_values) %>%
  rename(Field = c_fields, Operation = c_operations, Value = c_values)
kable(filters_table)

```
